models with mice

### 2024-03-10 withmice.Rmd
made lms with mice

@@

formice <- codedmodule2[, grep("State|age|urban_rural|gender|caste|religion|Solar|SOLAR|Nuclear|NUCLEAR|Coal|COAL|K_S|K_E|K_H|K_I|WEALTHLIM|MECHANISATION|DECISIONDECEN|INDUSTRYSMALL|ECONOMYLOCAL|ENVOVERDEV|OWNERPUB|OWNERREG|DECISIONCEN|INDUSTRYLARGE|ECONOMYGLOBAL|OWNERPVT|OWNERNOREG", names(codedmodule2))]%>%
        dplyr::select(-c("Language", "Risky_INDSolar", "Ben_INDSolar", "religiontext"))

@@

formice <- formice %>%
  mutate(across(Risky_Solar:age, ~factor(.x, ordered = TRUE)))%>%
  mutate(across(urban_rural:religion, ~factor(.x, ordered = FALSE)))%>%
  mutate(across(State, ~factor(.x, ordered = FALSE)))

str(formice)
summary(formice)

@@

## Mifa dor Kahan scale 
```{r, echo =FALSE}
library(mifa)
library(mice)
library(psych)

@@

kahanall <- codedmodule2 %>%
          dplyr::select(c(starts_with("K_")))

mikahan <- mifa(data= kahanall,
                cov_vars = everything(),
                n_pc= 2:4,
                ci = "fieller", 
                n_boot=1000,
                print = FALSE,
                progress= TRUE, 
                seed = 427,)

@@

@@

@@

#checking imputations: Kahan scale

```{r}
orgscores <- data.frame(kahanefascores)

summary(mikahan)
names(mikahan)

@@

```{r}
# Add an identifier column to each dataset
fct_scores$Dataset <- 'fct_scores'
orgscores$Dataset <- 'orgscores'

@@

ggplot(combined_data, aes(x = Dataset, y = MR1, fill = Dataset)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, aes(color = Dataset), size = 1.5, alpha = 0.5) +
  labs(title = "Boxplot of MR1 for Each Dataset", x = "", y = "MR1 Value") +
  scale_fill_manual(values = c("fct_scores" = "blue", "orgscores" = "red")) +
  scale_color_manual(values = c("fct_scores" = "blue", "orgscores" = "red")) +
  theme_minimal()

```
```{r}
ggplot(combined_data, aes(x = Dataset, y = MR2, fill = Dataset)) +
  geom_boxplot() +
  geom_jitter(width = 0.2, aes(color = Dataset), size = 1.5, alpha = 0.5) +
  labs(title = "Boxplot of MR2 for Each Dataset with Points", x = "", y = "MR2 Value") +
  scale_fill_manual(values = c("fct_scores" = "blue", "orgscores" = "red")) +
  scale_color_manual(values = c("fct_scores" = "blue", "orgscores" = "red")) +
  theme_minimal()

@@

ecopoln <- codedmodule2[, grep("NUCLEAR|WEALTHLIM|MECHANISATION|DECISIONDECEN|INDUSTRYSMALL|ECONOMYLOCAL|ENVOVERDEV|OWNERPUB|OWNERREG|DECISIONCEN|INDUSTRYLARGE|ECONOMYGLOBAL|OWNERPVT|OWNERNOREG", names(codedmodule2))]

miecopol <- mifa(data= ecopoln,
                cov_vars = everything(),
                n_pc= 2:4,
                ci = "fieller", 
                n_boot=1000,
                print = FALSE,
                progress= TRUE, 
                seed = 429,)

@@

@@

summary(complete_ecopol_5)
```

>>>>

# missing data pattern and MCAR

@@

formice1 <- codedmodule2[, grep("State|age|urban_rural|gender|caste|religion|Solar|SOLAR|Nuclear|NUCLEAR|Coal|COAL|K_S|K_E|K_H|K_I|WEALTHLIM|MECHANISATION|DECISIONDECEN|INDUSTRYSMALL|ECONOMYLOCAL|ENVOVERDEV|OWNERPUB|OWNERREG|DECISIONCEN|INDUSTRYLARGE|ECONOMYGLOBAL|OWNERPVT|OWNERNOREG", names(codedmodule2))]%>%
        dplyr::select(-c("Language", "Risky_INDSolar", "Ben_INDSolar", "religiontext")) %>%
          mutate(across(age, ~factor(.x, ordered = TRUE)))%>%  
          mutate(across(urban_rural:religion, ~factor(.x, ordered = FALSE)))%>%
          mutate(across(State, ~factor(.x, ordered = FALSE))) %>%
          mutate(Uppercaste = ifelse(caste %in% c('Brahmin', 'General'), 1, 0)) %>%
          mutate(Male = ifelse(gender == 'Male', 1, 0))%>%
          mutate(Hindu = ifelse(religion == 'Hinduism', 1, 0))%>%
           mutate(Age = case_when(
                                  age == "18-24 years old" ~ 1,
                                  age == "25-34 years old" ~ 2,
                                  age == "35-44 years old" ~ 3,
                                  age == "45-54 years old" ~ 4,
                                  age == "55-64 years old" ~ 5,
                                  age == "65-74 years old" ~ 6,
                                  age == "75 years or older" ~ 7))

@@

formice <- formice1 %>%
  mutate(across(Risky_Solar:age, ~factor(.x, ordered = TRUE)))%>%
  mutate(across(urban_rural:religion, ~factor(.x, ordered = FALSE)))%>%
  mutate(across(State, ~factor(.x, ordered = FALSE)))

#str(formice)
#summary(formice)

# litte's MCAR test

mcar_test(data=formice)

# p value = 1, which means MAR condition is met.

@@

# mice on entire dataset

#str(formice1)

#formice1.imp <- mice(formice1, m =5, maxit =5,  seed = 430, print = TRUE)

#formice1.imp$chainMean

#plot(formice1.imp)

# Diagnostic checking
```{r, echo= FALSE}

formice1.imp$imp$Risky_Nuclear

complete(formice1.imp)

library(mice)

# Select the variables you want to plot
selected_vars <- c("Risky_Nuclear", "Risky_Solar", "Risky_Coal")

# Use stripplot to plot only the selected variables

stripplot(formice1.imp, scales = list(x = list(relation = "free")),layout = c(3, 1), pch = 20, cex = 1.2)

# Density plot - entire dataset 
```{r, echo=FALSE}

densityplot(formice1.imp, scales = list(x = list(relation = "free")), layout = c(5, 1))

# propensity score check imputations

formice1.imp
#hc.na <- is.na(boys$hc)

#fit.hc <- with(imp.kendall, glm(hc.na ~ age + wgt + hgt + reg,
#family = binomial))
#ps <- rep(rowMeans(sapply(fit.hc$analyses, fitted.values)), 6)
```

# Checking MAR condition and missing data pattern for the Kahan Scale.
```{r, echo=FALSE}

kahanall <- codedmodule2 %>%
          dplyr::select(c(starts_with("K_"), "age", "gender", "caste", "State", "urban_rural", "religion")) %>%
          mutate(across(age:religion, ~factor(.x, ordered = FALSE)))

md.pattern(kahanall) #the variables with missing values: K_SHARM K_IPROTECT K_IPRIVACY K_SPROTECT K_SLIMCHOI K_IINTRFER K_HEQUAL K_HREVDIS1

mcar_test(data=kahanall) # p values is less than 0.05 - the condition for MCAR is not met

# dataset with missing variables to compare and test group comaprisons

MARkahan <- codedmodule2 %>%
          dplyr::select(c("K_SHARM", "K_IPROTECT", "K_IPRIVACY", "K_SPROTECT", "K_SLIMCHOI", "K_IINTRFER",  "K_HEQUAL", "K_HREVDIS1" , "age", "gender", "caste", "State", "urban_rural", "religion"))

MARkahan$K_SHARM_MAR <- MARkahan$K_SHARM

MARkahan$K_IPROTECT_MAR <- MARkahan$K_IPROTECT

MARkahan$K_SPROTECT_MAR <- MARkahan$K_SPROTECT

MARkahan$K_SLIMCHOI_MAR <- MARkahan$K_SLIMCHOI

library(finalfit)

# "K_SHARM" = "(C)Sometimes the government needs to make laws that keep people from hurting themselves.",
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "K_SHARM_MAR"
MARkahan %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the K_SHARM variable.")

#   "K_IPROTECT" = "(I)It’s not the government’s business to try to protect people from themselves."
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "K_IPROTECT_MAR"
MARkahan %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the K_IPROTECT variable.")

#  "K_SPROTECT" = "(C)The government should do more to advance society’s goals, even if that means limiting the freedom and choices of individuals.",
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "K_SPROTECT_MAR"
MARkahan %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the K_SPROTECT variable.")

#  "K_SLIMCHOI" = "(C)The government should put limits on the choices individuals can make so they don’t get in the way of what’s good for society.",
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "K_SLIMCHOI_MAR"
MARkahan %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the K_SPROTECT variable.")

## Mifa for Kahan scale 
```{r, echo =FALSE}

library(mifa)
library(mice)
library(psych)

@@

mikahan <- mifa(data= kahanall,
                cov_vars  = -c("age", "gender", "caste", "State", "urban_rural", "religion"),
                m = 5,
                n_pc= 2:4,
                ci = "fieller", 
                n_boot=1000,
                print = FALSE,
                progress= TRUE, 
                seed = 427,)

@@

# checking merging of imputations  : kahan scale

kahan_imp <- mice(kahanall, m =5, maxit =30,  seed = 427, print = TRUE)

kahan_imp$chainMean

@@

plot(kahan_imp)

@@

#checking imputations: Kahan scale - are values plausible ?

yes all value are whole numbers between 1 and 5

summary(mikahan)

str(data_imp)
names(mikahan)

@@

```{r, echo = False }

densityplot(kahan_imp, scales = list(x = list(relation = "free")), layout = c(5, 1))

```{r, echo=FALSE}
# Add an identifier column to each dataset
fct_scores$Dataset <- 'impscores'
orgscores$Dataset <- 'orgscores'

@@

ggplot(combined_data, aes(x = MR1, color = Dataset)) +
  geom_density(fill = NA, alpha = 0.5) +
  labs(title = "Density Plot of Hierachy- Egalitarianism for Each Dataset", x = "HE Value", y = "Density") +
  scale_color_manual(values = c("impscores" = "blue", "orgscores" = "red")) +
  theme_minimal()

```{r}
ggplot(combined_data, aes(x = MR2, color = Dataset)) +
  geom_density(fill = NA, alpha = 0.5) +
  labs(title = "Density Plot of Communitarianism for Each Dataset", x = "Communitarianism Value", y = "Density") +
  scale_color_manual(values = c("impscores" = "blue", "orgscores" = "red")) +
  theme_minimal()

# ecopol scale - checking MAR

ecopoln <- formice1 %>%
  dplyr::select(
    DECISIONDECEN, DECISIONCEN, INDUSTRYSMALL, INDUSTRYLARGE,
    ECONOMYLOCAL, ECONOMYGLOBAL, ENVOVERDEV, OWNERPVT, OWNERPUB,
    OWNERREG, OWNERNOREG, WEALTHLIM, MECHANISATION,
    DISPLACENUCLEAR,  POLLUTENUCLEAR, HEALTHNUCLEAR, JOBSNUCLEAR, BEAUTYNUCLEAR,
    PRIDENUCLEAR, NPRIDENUCLEAR, DEVNUCLEAR, PROSPERNUCLEAR,
    RELYNUCLEAR,age, urban_rural, gender, caste, religion, State
  )

md.pattern(ecopoln) #the variables with missing values: PRIDENUCLEAR PROSPERNUCLEAR JOBSNUCLEAR HEALTHNUCLEAR BEAUTYNUCLEAR POLLUTENUCLEAR DISPLACENUCLEAR RELYNUCLEAR

mcar_test(data=ecopoln) # p values is less than 0.76 - the condition for MCAR is met

# dataset with missing variables to compare and test group comaprisons

MARecopol <- codedmodule2 %>%
          dplyr::select(c(DECISIONDECEN, DECISIONCEN, INDUSTRYSMALL, INDUSTRYLARGE,
    ECONOMYLOCAL, ECONOMYGLOBAL, ENVOVERDEV, OWNERPVT, OWNERPUB,
    OWNERREG, OWNERNOREG, WEALTHLIM, MECHANISATION,
    DISPLACENUCLEAR,  POLLUTENUCLEAR, HEALTHNUCLEAR, JOBSNUCLEAR, BEAUTYNUCLEAR,
    PRIDENUCLEAR, NPRIDENUCLEAR, DEVNUCLEAR, PROSPERNUCLEAR,
    RELYNUCLEAR,age, urban_rural, gender, caste, religion, State))

MARecopol$JOBSNUCLEAR_MAR <- MARecopol$JOBSNUCLEAR

MARecopol$PRIDENUCLEAR_MAR <- MARecopol$PRIDENUCLEAR

library(finalfit)

# JOBSNUCLEAR
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "JOBSNUCLEAR_MAR"
MARecopol %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the JOBSNUCLEAR variable.")

# PRIDENUCLEAR
explanatory = c("age", "gender", "caste", "State", "urban_rural", "religion")
dependent = "PRIDENUCLEAR_MAR"
MARecopol %>% 
  missing_compare(dependent, explanatory) %>% 
    knitr::kable(row.names=FALSE, align = c("l", "l", "r", "r", "r"), 
        caption = "comparisons between values of responders (Not missing) and 
        non-responders (Missing) on the PRIDENUCLEAR variable.")

@@

#ecopoln <- codedmodule2[, grep("NUCLEAR|WEALTHLIM|MECHANISATION|DECISIONDECEN|INDUSTRYSMALL|ECONOMYLOCAL|ENVOVERDEV|OWNERPUB|OWNERREG|DECISIONCEN|INDUSTRYLARGE|ECONOMYGLOBAL|OWNERPVT|OWNERNOREG|", names(codedmodule2))]

summary(formice1)

miecopol <- mifa(data= ecopoln,
                cov_vars  = -c("age", "gender", "caste", "State", "urban_rural", "religion"),
                n_pc= 2:4,
                ci = "fieller", 
                n_boot=1000,
                print = TRUE,
                progress= TRUE, 
                seed = 429,)

@@

imputations are pluasible- all natural numbers between 1 and 5

@@

summary(complete_ecopol_5)

# Density plots - 2 factors from eco-pol scale 
```{r}

nscores <- data.frame(ecopolall_scores$scores)

# Add an identifier column to each dataset
fct_scores2$Dataset <- 'imputed'
nscores$Dataset <- 'original'

# Combine the datasets for plotting
compare2 <- rbind(fct_scores2, nscores)

# Select only necessary columns (assuming 'Index' isn't needed anymore)
compare2<- compare2[, c("MR1", "MR2", "Dataset")]

library(ggplot2)

ggplot(compare2, aes(x = MR2, color = Dataset)) +
  geom_density(fill = NA, alpha = 0.5) +
  labs(title = "Density Plot of MR2 for Each Dataset", x = "MR2 Value", y = "Density") +
  scale_color_manual(values = c("imputed" = "blue", "original" = "red")) +
  theme_minimal()

# Binding imputed values to formice dataset

```{r, echo = FALSE}

# binding the imputed  factor analysis scores for kahan scale and eco-pol value scale to the larger dataset codedmodule2

ecopolnuclear <- fct_scores2 %>%
  rename(Ndevelop = MR1, Pdevelop = MR2)

kahannuclear <- fct_scores%>%
  rename(Hierarchy_Egalitarian = MR1, Communitarian = MR2)

formice1

allformice<- cbind(formice1,ecopolnuclear, kahannuclear)

#summary(formice)

#method_list <- rep("polr", ncol(formice))
#names(method_list) <- colnames(formice)

#skip_vars <- c("age",  "urban_rural", "gender",  "caste", "religion", "K_ERADEQ2","K_ERADEQ1", "INDUSTRYSMALL","INDUSTRYLARGE","WEALTHLIM", "MECHANISATION", "State","K_IINTRFER",  "K_IPRIVACY",  "K_SHARM",  "K_IPROTECT",  "K_SLIMCHOI",  "K_SPROTECT",  "K_HEQUAL",  "K_HREVDIS1",  "K_EDISCRIM",    "K_EWEALTH",  "DECISIONDECEN",  "DECISIONCEN",     "ECONOMYLOCAL",  "ECONOMYGLOBAL",  "ENVOVERDEV",  "OWNERPVT",  "OWNERPUB",  "OWNERREG",  "DISPLACESOLAR",  "POLLUTESOLAR",  "HEALTHSOLAR",  "JOBSSOLAR",  "BEAUTYSOLAR",  "PRIDESOLAR",  "NPRIDESOLAR",  "DEVSOLAR",  "PROSPERSOLAR",  "RELYSOLAR",  "DISPLACECOAL",  "POLLUTECOAL",  "HEALTHCOAL",  "JOBSCOAL",  "BEAUTYCOAL",  "PRIDECOAL",  "NPRIDECOAL",  "DEVCOAL",  "PROSPERCOAL",  "RELYCOAL",  "DISPLACENUCLEAR",  "POLLUTENUCLEAR",  "HEALTHNUCLEAR",  "JOBSNUCLEAR",  "BEAUTYNUCLEAR",  "PRIDENUCLEAR",  "NPRIDENUCLEAR",  "DEVNUCLEAR",  "PROSPERNUCLEAR",  "RELYNUCLEAR", "OWNERNOREG")

#Set method to "" for variables to skip
#method_list[skip_vars] <- ""

#nuclear_imp1 <- mice(formice, method=method_list, maxit = 5, m =5,  seed = 123, print = TRUE)

print(nuclear_imp1)
summary(nuclear_imp1)
nuclear_imp1$imp$Risky_Nuclear

str(allformice)
library(mice)
library(dplyr)

# Initialize the mice process to create the predictor matrix
ini <- mice(allformice, maxit = 0, pri = FALSE)

# Extract the predictor matrix from the initialized object
pred <- ini$predictorMatrix

# Set the columns for "Uppercaste", "Male", "Hindu", and "Age" to 0
pred[, "Uppercaste"] <- 0
pred[, "Male"] <- 0
pred[, "Hindu"] <- 0
pred[, "Age"] <- 0
pred[, "Ndevelop"] <- 0
pred[, "Pdevelop"] <- 0
pred[, "Hierarchy_Egalitarian"] <- 0
pred[, "Communitarian"] <- 0
# Identify columns that start with "K_" and set them to 0
k_columns <- grep("^K_", colnames(formice1))
pred[, k_columns] <- 0

# Perform the mice imputation with the updated predictor matrix
allformice.imp <- mice(allformice, m = 5, maxit = 10, predictorMatrix = pred, seed = 430, print = TRUE)

summary(allformice.imp)
allformice.imp$chainMean

plot(allformice.imp)

nuclearfit1 <- with(allformice.imp, lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age))

nuclearfit2 <- with(allformice.imp, lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State))

nuclearfit3 <- with(allformice.imp, lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State+ Communitarian + Hierarchy_Egalitarian))

nuclearfit4 <- with(allformice.imp, lm(Risky_Nuclear ~ Uppercaste + Male + Hindu + urban_rural + Age + State+ Communitarian + Hierarchy_Egalitarian + Pdevelop + Ndevelop))

summary(pool(nuclearfit1))
summary(pool(nuclearfit2))
summary(pool(nuclearfit3))
summary(pool(nuclearfit4))

# LM with demographic variables

# The pool() function from the mice package creates a list containing the results,
# so we need to access the summary of the model directly.

pooled_fit <- pool(nuclearfit1)

pooled_summary <- summary(pooled_fit)

coefficients_df <- pooled_summary
# To ensure the numeric values are rounded to two decimal places, you can use:
coefficients_df <- round(coefficients_df[, -1], 2)  # Skipping the first column as it's factors (term names)
coefficients_df$term <- pooled_summary$term  # Adding back the unrounded terms

# Print the dataframe
data.frame(coefficients_df) %>%
  dplyr::select(term, estimate, std.error, statistic, df, p.value)
```

# LM with demographic variables and State

# The pool() function from the mice package creates a list containing the results,
# so we need to access the summary of the model directly.

pooled_fit2 <- pool(nuclearfit2)

pooled_summary2 <- summary(pooled_fit2)

coefficients_df2 <- pooled_summary2
# To ensure the numeric values are rounded to two decimal places, you can use:
coefficients_df2 <- round(coefficients_df2[, -1], 2)  # Skipping the first column as it's factors (term names)
coefficients_df2$term <- pooled_summary2$term  # Adding back the unrounded terms

# Print the dataframe
data.frame(coefficients_df2) %>%
  dplyr::select(term, estimate, std.error, statistic, df, p.value)
```

# LM with demographic variables and State

```{r}
# The pool() function from the mice package creates a list containing the results,
# so we need to access the summary of the model directly.

pooled_fit3 <- pool(nuclearfit3)

pooled_summary3 <- summary(pooled_fit3)

coefficients_df3 <- pooled_summary3
# To ensure the numeric values are rounded to two decimal places, you can use:
coefficients_df3 <- round(coefficients_df3[, -1], 2)  # Skipping the first column as it's factors (term names)
coefficients_df3$term <- pooled_summary3$term  # Adding back the unrounded terms

# Print the dataframe
data.frame(coefficients_df3) %>%
  dplyr::select(term, estimate, std.error, statistic, df, p.value)
```

# LM with demographic variables and State

# The pool() function from the mice package creates a list containing the results,
# so we need to access the summary of the model directly.

pooled_fit4 <- pool(nuclearfit4)

pooled_summary4 <- summary(pooled_fit4)

coefficients_df4 <- pooled_summary4
# To ensure the numeric values are rounded to two decimal places, you can use:
coefficients_df4 <- round(coefficients_df4[, -1], 2)  # Skipping the first column as it's factors (term names)
coefficients_df4$term <- pooled_summary4$term  # Adding back the unrounded terms

# Print the dataframe
data.frame(coefficients_df4) %>%
  dplyr::select(term, estimate, std.error, statistic, df, p.value)
```
